Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-06-08 manuscript/manuscript.Rnw
Housekeeping genes shape in different species

%\documentclass[english,12pt]{article}
\documentclass[doublespacing]{bmcart_dla}
\usepackage{fontspec}
\setmainfont{FreeSans}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSans}
\setmathsf{FreeSans}
\setmathtt{FreeSans}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
\usepackage{url}
% \usepackage{fancyhdr}
\usepackage{graphicx}
% \usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{sansmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
% \usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
\usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
\renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
% \newenvironment{narrow}[2]{%
%   \begin{list}{}{%
%       \setlength{\topsep}{0pt}%
%       \setlength{\leftmargin}{#1}%
%       \setlength{\rightmargin}{#2}%
%       \setlength{\listparindent}{\parindent}%
%       \setlength{\itemindent}{\parindent}%
%       \setlength{\parsep}{\parskip}}%
%   \item[]}{\end{list}}
% \def\newblock{\hskip}
% \newenvironment{paperquote}{%
%   \begin{quote}%
%      \it
%   }%
%   {\end{quote}}
% \renewcommand{\textfraction}{0.15}
% \renewcommand{\topfraction}{0.85}
% \renewcommand{\bottomfraction}{0.65}
% \renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
% \newenvironment{enumerate*}%
%   {\begin{enumerate}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{enumerate}}
% \newenvironment{itemize*}%
%   {\begin{itemize}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{itemize}}
% \setcounter{tocdepth}{3}
% \oddsidemargin 0.0in 
% \textwidth 6in
% \raggedbottom
% \clubpenalty = 10000
% \widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
% \renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

@@

>>>>

%\documentclass[english,12pt]{article}
\documentclass[doublespacing]{bmcart_dla}
\usepackage{fontspec}
\setmainfont{FreeSans}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSans}
\setmathsf{FreeSans}
\setmathtt{FreeSans}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
\usepackage{url}
% \usepackage{fancyhdr}
\usepackage{graphicx}
% \usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{sansmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
% \usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
\usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
\renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
% \newenvironment{narrow}[2]{%
%   \begin{list}{}{%
%       \setlength{\topsep}{0pt}%
%       \setlength{\leftmargin}{#1}%
%       \setlength{\rightmargin}{#2}%
%       \setlength{\listparindent}{\parindent}%
%       \setlength{\itemindent}{\parindent}%
%       \setlength{\parsep}{\parskip}}%
%   \item[]}{\end{list}}
% \def\newblock{\hskip}
% \newenvironment{paperquote}{%
%   \begin{quote}%
%      \it
%   }%
%   {\end{quote}}
% \renewcommand{\textfraction}{0.15}
% \renewcommand{\topfraction}{0.85}
% \renewcommand{\bottomfraction}{0.65}
% \renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
% \newenvironment{enumerate*}%
%   {\begin{enumerate}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{enumerate}}
% \newenvironment{itemize*}%
%   {\begin{itemize}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{itemize}}
% \setcounter{tocdepth}{3}
% \oddsidemargin 0.0in 
% \textwidth 6in
% \raggedbottom
% \clubpenalty = 10000
% \widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
% \renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

@@

In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. However, there is still a large
difference between the set of housekeeping genes which are expressed
in \human and those which are expressed in all placentas.

### 2016-06-08 manuscript/supplemental_material.Rnw
Housekeeping genes shape in different species

>>>>

\begin{figure}
<<housekeeping.placenta.distribution,fig.width=8,fig.height=8,out.height="0.8\\textheight",echo=FALSE,message=FALSE>>=
## housekeeping set as apposed to non-housekeeping set
cptg.h <- core.placenta.transcriptome.genes.shape[(human_name %in% housekeeping.genes.superset[,gene_short_name])]

cpt.h.distribution.shape <- 
    lapply(c("all_expression",
             "all_but_one_expression",
             "all_but_two_expression",
             "all_but_three_expression",
             "all_but_four_expression"),
           function(x){
               calculate.distribution.shape(cptg.h[,x,with=FALSE][[1]])[,`Minimum Expression in`:=x]
           })
fix.expression.names <- function(x){
    x <- gsub("^all_expression$",
              paste0(num.species," of ",num.species," species"),
              x
              )
    nc.levels <- paste0(num.species," of ",num.species," species")
    ## number coutner
    nc <- c("one"=1,"two"=2,"three"=3,"four"=4)
    for (a in names(nc)) {
        x <- gsub(paste0("^all_but_",a,"_expression$"),
              paste0(num.species-nc[a]," of ",num.species," species"),
              x
              )
        nc.levels <-
            c(nc.levels,
              paste0(num.species-nc[a]," of ",num.species," species"))
    }
    return(factor(x,levels=nc.levels))
}

cpt.h.distribution.shape <- 
    rbindlist(cpt.h.distribution.shape)

cpt.h.distribution.shape <- 
    cpt.h.distribution.shape[,`Minimum Expression in`:=
                                fix.expression.names(`Minimum Expression in`)]

setnames(cpt.h.distribution.shape,
         "FPKM",
         "Expression (FPKM)")

print(ggplot(cpt.h.distribution.shape[`Expression (FPKM)` > 0.1,],
             aes(x=`Expression (FPKM)`,color=`Minimum Expression in`,y=num_genes))
      + geom_line(lwd=1)
      + xlab("Minimum Expression cutoff (FPKM)")
      + ylab("Genes with N-th Minimum Expression")
      + scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                      labels = trans_format("log10", math_format(10^.x)))
      + scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                      labels = trans_format("log10", math_format(10^.x)))
      + geom_vline(xintercept=10)
      + geom_segment(mapping=aes(y=V1,
                         yend=V1,
                         xend=10,
                         x=250,
                         color=`Minimum Expression in`),
                     data=cpt.h.distribution.shape[`Expression (FPKM)` > 10,
                                              max(num_genes),`Minimum Expression in`],
                     arrow=arrow(length=unit(0.5,"cm"))
                   )
      + geom_text(mapping=aes(x=250,
                      y=V1,
                      label=paste0(V1," genes"),override=TRUE),
                  data=cpt.h.distribution.shape[`Expression (FPKM)` > 10,
                      max(num_genes),`Minimum Expression in`],
                  show_guide=FALSE,
                  hjust=0
                  )
      + guides(color=guide_legend(title="Minimum\nExpression in"))
      )
@ 
\caption{Distribution shape of housekeeping genes in placentas out of
  \Sexpr{pts[,sum(expressed.housekeeping)]} \human housekeeping 
  genes which are expressed in any placenta. The vertical axis
  shows the number of 1:1 orthologous \human housekeeping genes
  expressed in 
  placenta transcriptomes if %
  \Sexpr{cpt.h.distribution.shape[,levels(`Minimum Expression in`)[1]]}
  %
  (or %
  \Sexpr{cpt.h.distribution.shape[,levels(`Minimum Expression in`)[2]]},
  % 
  etc.) have a minimum expression in FPKM equal to the horizontal axis.
  For example, if the cutoff is set to 10 (vertical black line), 
  % 
  \Sexpr{cpt.h.distribution.shape[`Expression (FPKM)` > 10 
    & `Minimum Expression in` == '14 of 14 species',max(num_genes)]}
  %
  genes would be included if the FPKM of a set of 1:1 orthologs of all 14 species must be at
  least 10, %
  \Sexpr{cpt.h.distribution.shape[`Expression (FPKM)` > 10 
    & `Minimum Expression in` == "13 of 14 species",max(num_genes)]}
  %
  1:1 orthologs would be included if 13 of 14 species have an
   with an FPKM of at least 10, and 
  \Sexpr{cpt.h.distribution.shape[`Expression (FPKM)` > 10 
    & `Minimum Expression in` == '10 of 14 species',max(num_genes)]}
  genes would be included if 10 of 14 species
  must have an FPKM of at least 10.}
\label{fig:core_housekeeping_transcriptome_shape}
\end{figure}
